Constraining gravitational wave amplitude birefringence with GWTC-3

### 2023-04-06 src/tex/ms.tex
edits to comparison with Masha's paper

\citet{Okounkova_2022} gave a constraint on frequency-independent \ac{GW} amplitude birefringence by performing statistical analysis on GWTC-2.
We take $f=100\, \mathrm{Hz}$ as the reference frequency to make the comparison.
They were able to constrain $\tilde{\kappa}$ at $d_C=d_C(z=0.3)$ to be $|\tilde{\kappa}| \lesssim 0.74$ at 1 $\sigma$, where
\begin{equation}
    \tilde{\kappa}d_C = \left[{\kappa}\frac{d_C}{1\, \mathrm{Gpc}}\frac{f}{100\, \mathrm{Hz}}\right]_{f=100\, \mathrm{Hz}}\,.
\end{equation}
Note that $\tilde{\kappa}$ is in units of $1/\mathrm{Gpc}$, while $\kappa$ is dimensionless.
Equivalently, their constraint is $|\kappa| \lesssim 0.74$ at 1 $\sigma$.
We obtained a tighter $68.27\%$ upper limit of $|\kappa| <$ \variable{output/restricted_absolute_kappa_68.txt}.

>>>>

Our primary constraint on birefringence is given by the GWTC-3 combined constraint on the birefringent opacity, $\kappa =$ \variable{output/restricted_kappa_median.txt} with 90\% credibility, obtained in Sec.~\ref{sec:results:gwtc} (blue band in Fig.~\ref{fig:violin_kappa}).
This can be compared to previous measurements in the literature.

\citet{Okounkova_2022} produced a constraint on frequency-independent \ac{GW} amplitude birefringence from the distribution of measured inclinations of \acp{BBH} in GWTC-2.
As opposed to our Eq.~\eqref{eq:waveform_modification}, this reference parameterized the effect of birefringence as
\begin{equation}
    h_{L/R}^{\mathrm{br}}(f) =
    h_{L/R}^{\mathrm{GR}}(f) \times
    \exp\left(\pm \tilde{\kappa}\, d_C\right)\,,
\end{equation}
where $\tilde{\kappa}$ is a frequency-independent opacity (simply denoted ``$\kappa$'' in Ref.~\cite{Okounkova_2022}).
Assuming a fiducial redshift of $z=0.3$ for all events in GWTC-2, \citet{Okounkova_2022} obtained a constraint of $|\tilde{\kappa}| \lesssim 0.74\, {\rm Gpc}^{-1}$ at 68\% credibility.
Taking a reference frequency of $f=100\, \mathrm{Hz}$, which is representative of LIGO-Virgo events, we may compare that constraint to ours by noting that
\begin{equation}
    \tilde{\kappa}d_C = \left[{\kappa}\frac{d_C}{1\, \mathrm{Gpc}}\frac{f}{100\, \mathrm{Hz}}\right]_{f=100\, \mathrm{Hz}}\,.
\end{equation}
per Eq.~\eqref{eq:waveform_modification}, and hence $\tilde{\kappa} = \kappa / {\rm Gpc}$ at $f=100\, {\rm Hz}$.
In terms of this fiducial quantity, our constraint becomes $|\tilde{\kappa}| <$ \variable{output/restricted_absolute_kappa_68.txt} ${\rm Gpc}^{-1}$ at 68\% credibility.
